Visual Interrogation of Attention-Based Models for Natural Language Inference and Machine Comprehension

### 2018-08-20 EMNLP-demo-paper/NLPVis-demo-paper.tex
update camera ready

\section*{Acknowledgments}
This work was performed under the auspices of the U.S. Department of Energy by Lawrence Livermore National Laboratory under Contract DE-AC52-07NA27344. This work is also supported in part by NSF: CGV: Award:1314896, NSF:IIP Award :1602127 NSF:ACI:award 1649923, DOE/SciDAC DESC0007446, CCMSC DE-NA0002375, PIPER: ER26142 DE-SC0010498, and NVIDIA Corporation. This material is based upon work supported by the Department of Energy, National Nuclear Security Administration, under Award Number(s) DE-NA0002375.

>>>>

\section*{Acknowledgments}
This work was performed under the auspices of the U.S. Department of Energy by Lawrence Livermore National Laboratory under Contract DE-AC52-07NA27344. 
This work is also supported in part by NSF: CGV: Award: 1314896, NSF: IIP: Award: 1602127, NSF: ACI: award: 1649923, DOE/SciDAC DESC0007446, CCMSC DE-NA0002375, PIPER: ER26142 DE-SC0010498, and NVIDIA Corporation. This material is based upon work supported by the Department of Energy, National Nuclear Security Administration, under Award Number(s) DE-NA0002375.

### 2018-08-20 EMNLP-demo-paper/intro.tex
update camera ready

Deep neural networks have been successfully applied to various natural language processing tasks.
%where end-to-end training become a primary option.
As the design of neural networks evolves, there is a trend of increasing model complexity become clear in terms of both the architecture and the parameter count. 
%
Although expressive models help improve the prediction performance, the fundamental lack of interpretability leads many researchers to consider neural network models black boxes. 
%
The ability to interpret model internals and reason about predictions is essential for understanding the limitation of the models and improving upon them.

%Without understanding the background meaning that features carry, visualization relies
%on proper quantification analysis (cite). Such analysis is not easy to come up with,
%and is tied to specific task (cite).

@@

To address these challenges in interpreting attention-based models, 
we introduce a Python library that allows users to create web-based interactive visual analytics environments,
in which they can interrogate the model by perturbing the input text and observing the effects on the attention and prediction, or modifying the attention and studying how it affects predictions (see Figure~\ref{fig:modelPipeline}).
%
We demonstrate our library applied to two NLP tasks: natural language
inference (NLI) and machine comprehension (MC). 
%For the NLI task, we visualize
%decomposable attention networks which represents a series of strongly performing models \cite{Parikh2016}.
%For the MC task, we present your visualization methods on bidirectional attention flow model,
%which also represents a line of state-of-the-art models \cite{Seo2016}.

>>>>

Deep neural networks have been successfully applied to various natural language processing tasks.
%where end-to-end training become a primary option.
As the design of neural networks evolves, there is a clear trend of increasing model complexity in terms of both the architecture and the parameter count. 
%
Although expressive models help improve the prediction performance, the fundamental lack of interpretability leads many researchers to consider neural network models black boxes. 
%
The ability to interpret model internals and reason about predictions is essential for understanding the limitation of the models and improving upon them.

@@

To address these challenges in interpreting attention-based models, 
we introduce a Python library that allows users to create web-based interactive visual analytics environments,
in which they can interrogate the model by perturbing the input text and observing the effects on the attention and prediction, or modifying the attention and studying how it affects predictions (see Figure~\ref{fig:modelPipeline}).
%
We demonstrate our library applied to two NLP tasks: natural language
inference (NLI) and machine comprehension (MC).

### 2018-08-20 EMNLP-demo-paper/result.tex
update camera ready

\section{Applications}
We demonstrate the proposed visualization system on the decomposable attention network~\cite{parikh2016emnlp}
for the NLI task and the BIDAF model~\cite{Seo2016} for the MC task.

@@

\subsection{Machine Comprehension}
\label{sec:MCexample}
In the machine comprehension task, the goal is to select a span of text as the answer
to a question sentence.
%two sequences of texts are given: context and question.
%The goal is to select a span of text from the context that answers the question. 
%We provide a simple model formulation and refer the reader to the paper for details.
%%\begin{align}
%	$C^\prime, Q^\prime = BiLSTM(C), BiLSTM(Q)$,
%	$\overleftarrow{A} = u(C^\prime, Q^\prime)$,
%	$\overrightarrow{A} = v(C^\prime, Q^\prime)$,
%	$s = m(C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%	$e = n(s, C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%%\end{align}
%where $C$ and $Q$ are embedding matrices for the context and the question,
%$\overleftarrow{A}$ and $\overrightarrow{A}$ are bidirectional attention flows,
%$s$ and $e$ are probabilities for starting and ending indices of the answer span.
%Our system reveals the internal states of $\overleftarrow{A}$, $\overrightarrow{A}$,
%$s$ and $e$.
The attention information is encoded as a bidirectional alignment (i.e., from context to question and vice versa). 
%The key difference is how the question to context alignment is computed (see \cite{Seo2016} for more details).
Here, we refer the context to question attention as \emph{att1} and the question to context attention as \emph{att2} (which is a vector instead of a matrix according to \citet{Seo2016}). In this demo, we apply a min max normalization for \emph{att2} after the softmax layer to better distinguish different attention values.

>>>>

\section{Applications}
We demonstrate the proposed visualization system on the decomposable attention network~\cite{parikh2016emnlp}
for the NLI task and the BIDAF model~\cite{seo2016bidirectional} for the MC task.

@@

\subsection{Machine Comprehension}
\label{sec:MCexample}
In the machine comprehension task, the goal is to select a span of text as the answer
to a question sentence.
%two sequences of texts are given: context and question.
%The goal is to select a span of text from the context that answers the question. 
%We provide a simple model formulation and refer the reader to the paper for details.
%%\begin{align}
%	$C^\prime, Q^\prime = BiLSTM(C), BiLSTM(Q)$,
%	$\overleftarrow{A} = u(C^\prime, Q^\prime)$,
%	$\overrightarrow{A} = v(C^\prime, Q^\prime)$,
%	$s = m(C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%	$e = n(s, C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%%\end{align}
%where $C$ and $Q$ are embedding matrices for the context and the question,
%$\overleftarrow{A}$ and $\overrightarrow{A}$ are bidirectional attention flows,
%$s$ and $e$ are probabilities for starting and ending indices of the answer span.
%Our system reveals the internal states of $\overleftarrow{A}$, $\overrightarrow{A}$,
%$s$ and $e$.
The attention information is encoded as a bidirectional alignment (i.e., from context to question and vice versa). 
%The key difference is how the question to context alignment is computed (see \cite{Seo2016} for more details).
Here, we refer the context to question attention as \emph{att1} and the question to context attention as \emph{att2} (which is a vector instead of a matrix according to \citet{seo2016bidirectional}). In this demo, we apply a min max normalization for \emph{att2} after the softmax layer to better distinguish different attention values.
